---
title: "Prediction.Assignment.Writeup.Final"
author: "Atair Camargo"
date: "July, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
sessionInfo()
.libPaths("C:/Temp/Data.Science/Library")
```

## Prediction Assignment Writeup

### Background

  Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

### Data

  The training data for this project are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

  The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

### Downloading and assigning Data Files

    Base and test data sets are downloaded and loaded into two files.
    
```{r}
#Assign files' urls
plm.train.url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"

plm.test.url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"


#assign files names
train.file <- "pml-training.csv"

test.file <- "pml-testing.csv"

#download the files
if (!file.exists(train.file)){
    download.file(plm.train.url, train.file)
}

if (!file.exists(test.file)){
    download.file(plm.test.url, test.file)
}

#load data into tables
plm.train.data <- read.csv(file = "pml-training.csv", header = TRUE, sep = ",", na.strings=c("#DIV/0!"," ", "", "NA", "NAs", "NULL"))
plm.test.data <- read.csv(file = "pml-testing.csv", header = TRUE, sep = ",", na.strings=c("#DIV/0!"," ", "", "NA", "NAs", "NULL"))
dim(plm.train.data)
dim(plm.test.data)
```
### Data Cleaning
  Visualizing the data, it can be seen several NAs, so NAs will be removed from Data.Frame.

```{r}
# Remove NAs
clean.plm.train.data <- plm.train.data[, which(colSums(is.na(plm.train.data)) == 0)]

clean.plm.test.data <- plm.test.data[, which(colSums(is.na(plm.test.data)) == 0)]
#Checking Reduction
dim(clean.plm.train.data)
dim(clean.plm.test.data)
```
### Data Transformation
  It can be seen that the first 7 columns are not relevant to prediction, so they will be removed.
```{r}
#Remove columns
final.plm.train.data <- clean.plm.train.data[,-c(1:7)]
final.plm.test.data <- clean.plm.test.data[,-c(1:7)]
dim(final.plm.train.data)
dim(final.plm.test.data)
```
### Data Slicing
  Data will be split in order to perform validation.
  
```{r warning=FALSE}
library(caret)
#Split 60% of the data
plm.inTrain <- createDataPartition(y=final.plm.train.data$classe, p=0.60, list=FALSE )

plm.train <- final.plm.train.data[plm.inTrain,]

plm.valid <- final.plm.train.data[-plm.inTrain,]

#Check data sets dimensions
dim(plm.train)
dim(plm.valid)
dim(final.plm.test.data)
```
### Define Pre Process Method
  Cross Validation and Principal Components Analysis
  
```{r}
#Define Pre Process Method
library("caret")
control <- trainControl(method="cv", number=3, p = 0.75, classProbs = TRUE, preProcOptions= "pca", allowParallel=TRUE)
``` 
### Building Models
  Modelling with Random Forest.
    Checking hardware performance as well.

```{r cache=TRUE, warning=FALSE}
# Random Forest
set.seed(2018)
system.time(fit.td.rf <- train(classe~., data=plm.train, method="rf", metric = "ROC", maximize = TRUE, trControl=control))
```

  Modelling with boosted trees "GBM".

```{r cache=TRUE, warning=FALSE}
# GBM
set.seed(2018)
system.time(fit.td.gbm <- train(classe~., data=plm.train, method="gbm", metric = "ROC", maximize = TRUE, trControl=control, verbose = FALSE))
```



```{r cache=TRUE, warning=FALSE}
# summarizing results
set.seed(2018)
results <- resamples(list(rf=fit.td.rf, gbm=fit.td.gbm))
# Comparison Table
summary(results)
```




```{r}
# boxplot comparison
bwplot(results)
# Dot-plot comparison
dotplot(results)
```


```{r}
pred1 <- predict(fit.td.rf, plm.valid)
confusionMatrix(plm.valid$classe, pred1)
```



```{r}
pred2 <- predict(fit.td.gbm, plm.valid)
confusionMatrix(plm.valid$classe, pred2)
```
### Result
  Results show that the random forest model has the highest accuracy in cross validation. Therefore, it will be used for prediction.

### Prediction
Random forest model for prediction

```{r}

pred <- predict(fit.td.rf, newdata=final.plm.test.data)
pred
````



